## Introduction
This report presents the results of the analysis on diamond data.

## Analysis

```{r load-dataset, echo=TRUE} # nolint
# Load the dataset
data <- read.csv("C:/Users/lenovo/Documents/Projects/Data Analysis/DiamondDataComplete.csv")

# Data quality check
summary(data)
str(data)

# Check for missing values
colSums(is.na(data))

# Boxplot of price by cut
library(ggplot2)
ggplot(data, aes(x = cut, y = price)) + 
  geom_boxplot() + 
  ggtitle("Boxplot of Price by Cut")

# Scatterplot of price vs carat with linear fit
ggplot(data, aes(x = carat, y = price)) + 
  geom_point() + 
  geom_smooth(method = "lm") +
  ggtitle("Scatterplot of Price vs Carat with Linear Fit")

library(caret)
set.seed(123)
trainIndex <- createDataPartition(data$price, p = 0.7, list = FALSE)
trainData <- data[trainIndex, ]
testData <- data[-trainIndex, ]

# Linear regression model
model <- lm(price ~ ., data = trainData)
summary(model)

### Predictions and Evaluation
library(Metrics)
```{r predictions-evaluation, echo=TRUE, results='markup'} # nolint
# Predictions
tryCatch({
  predictions <- predict(model, newdata = testData)
  
  # Alternative RMSE calculation
  rmse_value <- rmse(predictions, testData$price)
  
  # Correlation between predictions and actual values
  correlation <- cor(predictions, testData$price)
  
  # Print results
  print("Evaluation Metrics:")
  print(paste("RMSE: ", rmse_value))
  print(paste("Correlation: ", correlation))
}, error = function(e) {
  print("Error during prediction or evaluation:")
  print(e)
})

# Normalize data and perform KNN
library(class)
normData <- scale(data[, sapply(data, is.numeric)])
knn_result <- knn(
  train = normData[trainIndex, ],
  test = normData[-trainIndex, ],
  cl = data$cut[trainIndex],
  k = 5
)
knn_result[1:5]

# Train and predict using C5.0 model
library(C50)
trainData$cut <- as.factor(trainData$cut)
testData$cut <- as.factor(testData$cut)
c50_model <- C5.0(trainData[, -which(names(trainData) == "cut")], trainData$cut)
c50_pred <- predict(c50_model, testData)
head(c50_pred)

# Train and predict using ANN
library(nnet)
ann_model <- nnet(cut ~ ., data = trainData, size = 10, linout = FALSE)
ann_pred <- predict(ann_model, testData, type = "class")
head(ann_pred)